Adaptive group-regularized logistic elastic net regression

### 2019-04-02 docs/manuscript.Rnw
small changes to text of manuscript.Rnw and supplement.Rnw

% objects
\newcommand{\x}{\mathbf{x}}
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\m}{\mathbf{m}}

% functions and operators
\newcommand{\tr}{^{\text{T}}}
\newcommand{\expit}{\text{expit}}
\newcommand{\argmax}{\text{argmax} \,}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\diag}{\text{diag}}
\newcommand{\trace}{\text{tr}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

@@

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features 
		is often available. Examples of such external information in omics research
		are: (a) $p$-values from a previous study, (b) a summary of prior 
		information, and (c) omics annotation. The inclusion of this information in
		the analysis may enhance classification performance and feature selection,
		but is not straightforward in the standard regression setting. As a solution
		to this problem, we propose a group-regularized (logistic) elastic net
		regression method, where each penalty parameter corresponds to a group of
		features based on the external information. The method, termed 
		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
		regression to estimate both the model and penalty parameters in an 
		approximate empirical-variational Bayes framework. Simulations and
		applications in two different molecular marker and cancer type studies show
		that, if the partitioning of the features is informative, classification 
		performance and feature
		selection are indeed enhanced.}
	\end{abstract}

@@

	Two such studies are introduced in Sections \ref{sec:colorectal} and
	\ref{sec:oral}. The first study is concerned with treatment 
	response prediction for colorectal cancer patients, based on sequenced
	microRNAs. In the second study we attempt to predict lymph node metastases,
	using RNAseq data. In both settings additional information on the 
	features is available in the form of a partitioning. Furthermore, in both
	examples data is sparse, so any extra included information may help to boost
	predictive performance. Inclusion of such feature partitions in 
	the analysis is not straightforward with standard methods.

	At first sight, such an \textit{a priori} grouping of the features suggests 
	the group lasso
	\cite[]{meier_group_2008, yuan_model_2006} or one of its extensions, because
	the group lasso penalises and selects features at the group level. 
	There are two limitations to this approach: the group lasso (a) selects entire
	groups instead of single 
	features and (b) does not penalise adaptively: all groups are penalised 
	equally. Extensions such as the sparse group lasso 
	\cite[]{simon_sparse-group_2013} deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is,
	each group of variables receives its own penalty parameter. An apparent issue
	with this differential penalization is the estimation of the penalty 
	parameters. Naive estimation may be done by cross-validation. However, 
	cross-validation requires re-estimation of the model over a grid of penalty
	parameters. The size of this grid increases exponentially with the number of
	penalty parameters. Consequently, it quickly becomes computationally 
	infeasible. We therefore propose an efficient alternative: empirical Bayes
	estimation of the penalty parameters, which corresponds to hyperparameter
	estimation in the Bayesian prior framework. Because of the ubiquity of binary
	outcome data in omics research, we focus on the logistic elastic net.

@@

	Recently, \cite{zhang_novel_2019} introduced Variational Bayes Variable
	Selection (VBVS), a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
	adaptive inclusion of external information on the features.

@@

	Many forms of empirical Bayes exist, the most formal one being maximisation of
	the marginal likelihood with respect to the hyperparameters. The resulting 
	hyperparameter estimates are then plugged into the prior. The marginal 
	likelihood is often introduced as a measure of model evidence given the 
	observed data and is computed by integrating the product of likelihood and 
	prior with respect to the model parameters. In the case of the elastic net 
	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
	empirical Bayes posterior for $\bbeta$:
	\begin{align}
	  p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) 
	  \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = 
	  \frac{\int_{\btau} \L (\y ; \bbeta) 
	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
	  d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
	  \hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, 
	  \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
	  & = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta)
	  \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. 
	  \label{eq:mmlestimate}
	\end{align}
	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
	elastic net. In the omics setting the integrals are also high-dimensional, in 
	which case numerical and Monte Carlo approximation methods become tedious and 
	computationally expensive. Moreover, Laplace approximation is known to suffer 
	from low accuracy in many high-dimensional settings 
	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
	is described that estimates the hyperparameters. This EM algorithm iteratively
	maximises the expected joint log likelihood, such that the sequence:
	\begin{equation}\label{eq:ebEM}
	  \blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
	  \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
	\end{equation}
	converges to a local maximum of the marginal likelihood. The difficulty herein
	is in the calculation of the expected joint log likelihood. 
	\cite{casella_empirical_2001} suggests to approximate the expectation by its 
	Monte Carlo expectation. Although elegant and simple, this method requires a 
	converged MCMC sample from the posterior for every iteration: a 
	computationally intensive procedure. \cite{roy_selection_2017} introduce
	generalised importance sampling for the Bayesian elastic net, such that we
	just need a limited, pre-specified number of MCMC chains. However, this still
	requires several converged chains, so is not feasible in many high-dimensional
	omics settings. We propose to tackle this problem by approximating the
	expectation in (\ref{eq:ebEM}) using variational Bayes.

@@

	To simplify the computations of our variational approximation, we follow
	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
	\ref{sm-sec:polyagamma} of the SM for more 
	details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	  \bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of 
	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
	parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta_1 = \bomega$, $\btheta_2 = \bbeta$, $\btheta_3 = \btau$, and 
	$\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 \end{bmatrix}$, 
	calculus of variations gives the optimal distributions 
	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
	Kullback-Leibler divergence of the posterior to the approximate distribution 
	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
	renders both the posterior parameter calculations and the expected joint log 
	likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions in our variational Bayes implementation for the model parameters
	as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
	for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively. Naive calculation of the variational
	parameters is computationally expensive. In Section 
	\ref{sm-sec:computationalcomplexity} of the SM we show that 
	informed calculation of the parameters results in a significant reduction of 
	computational complexity.

@@

	We simulated data according to five different scenarios:
	\begin{enumerate}[(i)]
		\item \label{it:scenario1} differential signal between the groups and uniformly
		distributed model
		parameters,
		\item \label{it:scenario2} a large number of small groups of features,
		\item \label{it:scenario3} no differential signal between the groups, but 
		strong correlations 
		within groups of features,
		\item \label{it:scenario4} differential signal between the groups and 
		heavy-tailed distributed 
		model parameters, and
		\item \label{it:scenario5}a very sparse setting with no signal in some of 
		the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 
	\ref{sm-sec:simulations} of 
	the SM. In Scenario (\ref{it:scenario1}), the distribution of model 
	parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In Scenario (\ref{it:scenario2}), 
	which is similar to
	the simulations presented in the original group lasso paper 
	\cite[]{yuan_model_2006}, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate. The group lasso extensions are expected to perform better: they 
	estimate just one penalty parameter, so are less sensitive to overfitting.
	In Scenario (\ref{it:scenario3}), we expect all methods to underperform
	due to the high correlations between the features. The
	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
	performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	detect differential signal between
	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
	performance for all 
	methods, because of the sparse truth and lack of differential 
	signal.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples to compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
	from $p=900$ to $p=1000$ over the Scenarios. We varied the 
	covariance matrix of the features and simulation of 
	the model parameters $\beta_j$, $j=1, \dots, p$ between the
  Scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
	parametrisations). We 
	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation, we repeated each scenario 100 times 
	and report predictive performance measures 
	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
	For the methods with an $\alpha$ tuning parameter, we show the one that gives
	the best performance. Results for the other $\alpha$'s and additional
	performance measures are presented in Section 
	\ref{sm-sec:simulations} of the SM.

	In Scenario (\ref{it:scenario1})
	\texttt{gren} and \texttt{GRridge} don't suffer much from the wrong 
	distributional assumptions as they outperform the other methods.
	Scenario (\ref{it:scenario2}) follows our expectation \texttt{gren}, and to
	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
	the regular elastic net outperforms the group lasso methods, which are 
	designed to handle situations such as Scenario (\ref{it:scenario2}).
	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
	regular elastic net suffer more from the higher correlations. Here,
	the regular elastic net outperforms \texttt{gren}, which is an indication
	that the estimation of the penalty parameters is impaired by high
	correlation between features. Scenario (\ref{it:scenario4}) follows the 
	expected pattern, with the group-regularized methods (\texttt{GRridge} and
	\texttt{gren}) outperforming their respective non group-regularized 
	counterparts, as well as the sparse group lasso extensions.
	Surprisingly, in Scenario (\ref{it:scenario5}), \texttt{gren} outperforms all
	other methods, which is an indication that \texttt{gren} is able to pick up 
	very sparse differential signals between the groups.

<<lines_simulations_res1_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 1.", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res2_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 2.", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res3_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 3.", out.width="100%", fig.asp=1/2>>=
@

>>>>

% objects
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\ones}{\mathbf{1}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

@@

	\begin{abstract}
		{In high-dimensional data settings, additional information on the features 
		is often available. Examples of such external information in omics research
		are: (a) $p$-values from a previous study, (b) a summary of prior 
		information, and (c) omics annotation. The inclusion of this information in
		the analysis may enhance classification performance and feature selection,
		but is not straightforward in the standard regression setting. As a solution
		to this problem, we propose a group-regularized (logistic) elastic net
		regression method, where each penalty parameter corresponds to a group of
		features based on the external information. The method, termed 
		\texttt{gren}, makes use of the Bayesian formulation of logistic elastic net
		regression to estimate both the model and penalty parameters in an 
		approximate empirical-variational Bayes framework. Simulations and
		applications in two different cancer studies show
		that, if the partitioning of the features is informative, classification 
		performance and feature
		selection are indeed enhanced.}
	\end{abstract}

@@

	Two such studies are introduced in Sections \ref{sec:colorectal} and
	\ref{sec:oral}. The first study is concerned with treatment 
	response prediction for colorectal cancer patients, using sequenced
	microRNAs. In the second study we predict lymph node metastases in oral cancer
	patients, using RNAseq data. In both settings additional information on the 
	features is available in the form of a partitioning. Furthermore, in both
	studies data is sparse, so any extra included information may help to boost
	predictive performance. Inclusion of such extra information in 
	the analysis is not straightforward with standard methods.

	At first sight, such an \textit{a priori} grouping of the features suggests 
	the group lasso
	\cite[]{meier_group_2008, yuan_model_2006}, because
	the group lasso penalises and selects features at the group level. 
	There are two limitations to this approach: the group lasso (a) selects entire
	groups instead of single 
	features and (b) does not penalise adaptively: all groups are penalised 
	equally. Extensions such as the sparse group lasso 
	\cite[]{simon_sparse-group_2013} deal with (a), but do not address (b).

	A natural way to deal with (b) is through differential penalization: that is,
	each group of variables receives its own penalty parameter. An apparent issue
	with differential penalization is the estimation of the penalty 
	parameters. Naive estimation may be done by cross-validation. However, 
	cross-validation requires re-estimation of the model over a grid of penalty
	parameters. The size of this grid increases exponentially with the number of
	penalty parameters. Consequently, it quickly becomes computationally 
	infeasible. We therefore propose an efficient alternative: empirical Bayes
	estimation of the penalty parameters, which corresponds to hyperparameter
	estimation in the Bayesian prior framework. Because of the ubiquity of binary
	outcome data in omics research, we focus on the logistic elastic net.

@@

	Recently, \cite{zhang_novel_2019} introduced Variational Bayes Variable
	Selection (VBVS), a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features. 
	Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but these methods lack 
	adaptive inclusion of external information on the features.

@@

	Many forms of empirical Bayes exist, the most formal one being maximisation of
	the marginal likelihood with respect to the hyperparameters. The resulting 
	hyperparameter estimates are then plugged into the prior. The marginal 
	likelihood is often introduced as a measure of model evidence given the 
	observed data and is computed by integrating the product of likelihood and 
	prior with respect to the model parameters. In the case of the elastic net 
	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
	empirical Bayes posterior for $\bbeta$:
	\begin{align}
	  p_{\hat{\blambda}}(\bbeta | \y) & =
	  \frac{\int_{\btau} \L_{\hat{\blambda}}(\y, 
	  \bbeta, \btau) \, d\btau}{p_{\hat{\blambda}}(\y)} = 
	  \frac{\int_{\btau} \L (\y ; \bbeta) 
	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
	  d\btau}{p_{\hat{\blambda}}(\y)} \nonumber \\
	  \hat{\blambda} & := 
	  \underset{\blambda}{\argmax} p_{\blambda}(\y) = 
	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} 
	  \L_{\blambda}(\y, \bbeta, \btau) \, d\bbeta d\btau. \label{eq:mmlestimate}
	\end{align}
	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
	elastic net. In the omics setting the integrals are also high-dimensional, in 
	which case numerical and Monte Carlo approximation methods become tedious and 
	computationally expensive. Moreover, Laplace approximation is known to suffer 
	from low accuracy in many high-dimensional settings 
	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
	is described that estimates the hyperparameters. This EM algorithm iteratively
	maximises the expected joint log likelihood, such that the sequence:
	\begin{equation}\label{eq:ebEM}
	  \blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
	  \log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
	\end{equation}
	converges to a local maximum of the marginal likelihood. The difficulty herein
	is in the calculation of the expected joint log likelihood. 
	\cite{casella_empirical_2001} suggests to approximate the expectation by its 
	Monte Carlo expectation. Although elegant and simple, this method requires a 
	converged MCMC sample from the posterior for every iteration: a 
	computationally intensive procedure. \cite{roy_selection_2017} introduce
	generalised importance sampling for the Bayesian elastic net, such that we
	just need a limited, pre-specified number of MCMC chains. However, this still
	requires several converged chains, so is not feasible in many high-dimensional
	omics settings. We propose to tackle this problem by approximating the
	expectation in (\ref{eq:ebEM}) using variational Bayes.

@@

	To simplify the computations of our variational approximation, we follow
	\cite{polson_bayesian_2013} and introduce latent variables $\omega_i$, for 
	$i=1, \dots, n$. Conditional on $\bbeta$, the $\omega_i$ are independent of
	the $y_i$ and P\'{o}lya-Gamma distributed (see Section 
	\ref{sm-sec:polyagamma} of the SM for more 
	details). We augment Model (\ref{eq:enmodel2}) with:
	\begin{equation}\label{eq:logmodel}
	  \bomega | \bbeta \sim \prod_{i=1}^n \mathcal{PG}\(m_i, |\x_i \tr \bbeta| \).
	\end{equation}
	Our variational Bayes approximation to the posterior distribution of 
	(\ref{eq:enmodel2}) and (\ref{eq:logmodel}) factorizes over blocks of 
	parameters. We choose the blocks such that:
	\begin{equation}\label{eq:varbayesapproximation}
	  p (\bomega, \bbeta, \btau | \y) \approx Q(\bomega, \bbeta, \btau) 
	  = q_{\bomega} (\bomega) q_{\bbeta} (\bbeta) q_{\btau} (\btau).
	\end{equation}
	Writing $\btheta = \begin{bmatrix} \btheta_1 & \btheta_2 & \btheta_3 
	\end{bmatrix} = \begin{bmatrix} \bomega & \bbeta & \btau \end{bmatrix}$, 
	calculus of variations gives the optimal distributions 
	$q^*_{\btheta_j} (\btheta_j) \propto \exp \{\E_{\btheta \backslash \btheta_j} 
	[\log p (\btheta | \y)]\}$, where optimality is achieved in terms of the 
	Kullback-Leibler divergence of the posterior to the approximate distribution 
	\cite[]{neal_view_1998}. The approximation in (\ref{eq:varbayesapproximation})
	renders both the posterior parameter calculations and the expected joint log 
	likelihood as introduced in (\ref{eq:ebEM}) tractable.

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section \ref{sm-sec:variationaldistributions} 
	for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively. Naive calculation of the variational
	parameters is computationally expensive. In Section 
	\ref{sm-sec:computationalcomplexity} of the SM we show that 
	informed calculation results in a significant reduction of 
	computational complexity.

@@

	We simulated data according to five different scenarios:
	\begin{enumerate}[(i)]
		\item \label{it:scenario1} differential signal between the groups and uniformly
		distributed model
		parameters,
		\item \label{it:scenario2} a large number of small groups of features,
		\item \label{it:scenario3} no differential signal between the groups, but 
		strong correlations 
		within groups of features,
		\item \label{it:scenario4} differential signal between the groups and 
		heavy-tailed distributed 
		model parameters, and
		\item \label{it:scenario5}a very sparse setting with no signal in some of 
		the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 
	\ref{sm-sec:simulations} of 
	the SM. In Scenario (\ref{it:scenario1}), the distribution of model 
	parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In Scenario (\ref{it:scenario2}), 
	which is similar to
	the simulations presented in the original group lasso paper 
	\cite[]{yuan_model_2006}, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate. In Scenario (\ref{it:scenario3}), we expect all methods to 
	underperform due to the high correlations between the features. The
	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
	performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	detect differential signal between
	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
	performance for all methods, because of the sparse truth.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples to compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
	from $p=900$ to $p=1000$. The covariance matrix of the features and 
	simulation of the model parameters $\beta_j$, $j=1, \dots, p$ varies between
  Scenarios. We 
	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation, we repeated each scenario 100 times 
	and report average area under the receiver operator curve (AUC) and 
	Brier skill score in Figures \ref{fig:lines_simulations_res1_auc_briers}-
	\ref{fig:lines_simulations_res5_auc_briers}. AUC and Brier skill score are 
	measures of discrimination and calibration, respectively, and for both a 
	higher value indicates better predictive performance.
	For the methods with an $\alpha$ parameter, we display the best-performing 
	$\alpha$. Results for the other $\alpha$'s are presented in 
	Section \ref{sm-sec:simulations} of the SM.

	Overall, the ridge methods outperform the elastic net methods, which in turn
	outperform the group lasso methods, in terms of Brier skill score.
	In Scenario (\ref{it:scenario1})
	\texttt{gren} and \texttt{GRridge} outperform the other methods in terms
	of AUC.
	Scenario (\ref{it:scenario2}) the AUC follows our expectation: \texttt{gren}, 
	and to
	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
	the regular elastic net outperforms the group lasso methods.
	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
	regular elastic net suffer from the high correlations. The regular elastic net 
	outperforms \texttt{gren}, which is an indication that high correlations
	impair penalty parameter estimation. Scenario (\ref{it:scenario4}) follows the 
	expected pattern, with \texttt{GRridge} and
	\texttt{gren} outperforming their respective non group-regularized 
	counterparts, as well as the sparse group lasso extensions.
	Surprisingly, in Scenario (\ref{it:scenario5}), \texttt{gren} outperforms all
	other methods, which is an indication that \texttt{gren} is able to pick up 
	the sparse differential signal.

<<lines_simulations_res1_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario1}).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res2_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario2}).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res3_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario3}).", out.width="100%", fig.asp=1/2>>=
@

<<lines_simulations_res4_auc_briers, fig.cap="Average (a) AUC and (b) Brier skill score in simulation Scenario (\\ref{it:scenario4}).", out.width="100%", fig.asp=1/2>>=
@
